---
title: "A UN Report on the Relationship between GDP and Life Expectancy"
author: "Thea Bartlett"
date: "2024-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
```
# Introduction

This is a report and it was prepared for this workshop I'm in. Very important work. 
GDP is 100% absolutely correlated to life expectancy and this report will show that.
Dare I say it is evidence of causation!

```{r, include=FALSE}
gapminder_1997 <- read_csv("un-report/data/gapminder_1997.csv")
```
```{r gdp_lifeExp_1997_plot, echo=FALSE}
ggplot(data = gapminder_1997) +
  aes(x = gdpPercap, y = lifeExp, color = continent, size = pop/1e6) +
  geom_point() +
  labs(x= "GDP per cap", y = "Life Expectancy", title = "Do wealthy people live longer?",
       size = "Population (in millions)")
```

```{r, include=FALSE}
nCountries <- gapminder_1997 %>% 
  select(country) %>% 
  n_distinct()

minGDP <- gapminder_1997 %>% 
  summarize(round(min(gdpPercap))) %>% 
  pull()

maxGDP <- gapminder_1997 %>% 
  summarize(round(max(gdpPercap))) %>% 
  pull()
```
### Data Explanation

The above plot shows relationship between GDP per capita and life expectancy
for the total of `r nCountries` countries. The max GDP was `r maxGDP` and the min
was `r minGDP`.


|Summary of Data|
|----------|--------|
|Number of Countries|`r nCountries`|
|Min of GDP per capita|`r minGDP`|
|Max of GDP per capita|`r maxGDP`|

```{r, echo=FALSE}
#load library
library(knitr)

#print kable
gapminder_1997 %>% 
  filter(continent == "Oceania") %>% 
  kable()
```

